Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-08-01 decorrelation/decorrelation.tex
updates

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006}\needcite. These results show that quantum computation is feasible, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} , independance of errors \needcite , and even the magnitude of the noise \needcite . With these assumptions violated, promises about system performance cannot be made as easily.

@@

As examples of these deviations from the theory, there are many ways in which devices can behave in non-Markovian ways, such as drift in controls, \cite{Kelly2018} and general non-Markovian noise \cite{BlumeKohout2017}. Even if a given system behaves in a Markovian way, many models make simplifying assumptions about the noise on that system - such as positing that it has a simple structure. Because of its  classical simulability, the channel that is often chosen is the Pauli channel \needcite . From there, one can ask what models can honestly be represented by using these simple models \needcite, but the fundamental problem with this approach is the pessimism that comes along with it. As we get closer to having real, usable systems, we want models that not only give us guarantees, but faithfully represent our system.

Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to try to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown, potentially channel depdenent error, the realization has been made that coherent error in the process can be averaged away into incoherent error, at the cost of an additional compilation step. Some of these routines have even been demonstrated experiementally \cite{Ware2018}, and been shown to reduce the off diagonal elements in the Pauli Liouville representation \needcite of the channels.

A common feature of these routines is that take advantage of the fact that processes are generally enacted as a sequence of gates. By inserting a potentially expensive classical pre-compilation step, they can exploit sequence structure of a process to average away coherent error. In the case of \cite{Campbell2017}, the analysis takes the point of view of treating Pauli and Clifford gates as being free, which is out of reach of modern quantum computers\needcite. In \cite{Wallman2016} and the implementation in \cite{Ware2018}, the random frames are implemented in each sequence by precompiling the inserted Pauli frame rotations with the Clifford operations. In general, this cannot be done efficiently for non-Clifford operations, and there for this routine is demanding given the severe circuit depth constraints of modern quantum computers.\needcite

To remedy these problems, we take a different approach at the gate synthesis level. We propose to inject additional decorrelating randomness into the system through the use of \emph{balanced control solutions} (BCSs). BCSs are families of control solutions that all approximate the same target gate, but with balanced errors for any given instance of the noise Hamiltonian. That is, for a target gate, $U_T$, we seek a family of control solutions, $c_i(t)$, each implementing an approximation $U_i$ to the target gate, such that the family of unitary approximations is \emph{balanced}. A balanced family is one which satisfies, for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

@@

This techique can be used to turn coherent error into incoherent error and correlated error into uncorrelated error. The former gives the benefit that the error incurred from incoherent errors compound linearly, while coherent errors compound quadratically. The latter gives the benefit of reducing non-Markovian effects. This is a particularly useful property of BCSs, as diagnosing non-Markovianity in a gateset is challenging, but essential to perform reliable quantum computation. Many routines exist that can assess the quality of a gateset, however in the presence of non-Markovianity most of them become unreliable.

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome \cite{Merkel2013}, and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs change the error to a depolorizing channel, the correlations in the noise can be massively reduced, and the process can be made much closer to Markovian, at no real cost to the fidelity of the implemented gate. Unlike in \cite{Wallman2016, Campbell2017, Ware2018} this routine can be implemented natively in control electronics, and only requires that the onboard sequencer has the abiltiy to generate random numbers - no pre or post processing is required.

@@

\section{Experimental Results}\label{experimental}
To demonstrate the realizability of this routine, we implemented it on a superconducting qubit, to calibrate an $RX(\frac{\pi}{2})$ pulse. The qubit being used had a 20.5 $\mu$s T1, and a 11.00 $\mu$s T2, and the pulse shapes being considered were 30ns long top hat pulse sequences, generated via a set-and-hold signal generator using the amplitudes shown in \ref{fig:pulseshapes}. 
\begin{figure}[H]
  \centering
  \includegraphics[width=.75\columnwidth]{pulseshapes.png}
  \caption{The pulseshapes used in the routine. The original pulse was generated by iteratively applying $RX(\frac{\pi}{2})$ pulses, fitting the resulting sinusoid, and updating the controls. The variants were generated via optimal control, seeding the GRAPE algorithm with the initial pulse.}
  \label{fig:pulseshapes}
\end{figure}

>>>>

Steady progress has been made in the theory of quantum error correction, proving higher thresholds for increasingly general models of noise \cite{Aharonov2006}\needcite. These results show that quantum computation is feasible, however recent NISQ \cite{Preskill2018} devices have noise that is not only often above thresholds\needcite, but that also violates fundamental assumptions made by the models used in these results, such as Markovianity \cite{Kitaev1997} and independence of errors \needcite. With these assumptions violated, promises about system performance and correctness are difficult to make.

@@

There are many ways in which devices can behave in non-Markovian ways,\cite{Kelly2018, BlumeKohout2017}, and even if a system is Markovian, many models make simplifying assumptions about the noise on that system - such as positing that it has a simple structure. For instance Pauli channels are often chosen\needcite due to their classical simulability\needcite. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but scales poorly.

Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown, potentially channel depdenent error, coherent errors can be averaged away into incoherent error, at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing. Some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise.

%A common feature of these routines is that take advantage of the fact that processes are generally enacted as a sequence of gates. By inserting a potentially expensive classical pre-compilation step, they can exploit sequence structure of a process to average away coherent error. In the case of \cite{Campbell2017}, the analysis takes the point of view of treating Pauli and Clifford gates as being free, which is out of reach of modern quantum computers\needcite. In \cite{Wallman2016} and the implementation in \cite{Ware2018}, the random frames are implemented in each sequence by precompiling the inserted Pauli frame rotations with the Clifford operations. In general, this cannot be done efficiently for non-Clifford operations, and there for this routine is demanding given the severe circuit depth constraints of modern quantum computers.\needcite

In this Letter, we take a different approach to address these problems at the gate synthesis level. We propose to inject additional decorrelating randomness into the system through the use of \emph{balanced control solutions} (BCSs). BCSs are families of control solutions that all approximate the same target gate, but with balanced errors for any given instance of the noise Hamiltonian. That is, for a target gate, $U_T$, we seek a family of control solutions, $c_i(t)$, each implementing an approximation $U_i$ to the target gate, such that the family of unitary approximations is \emph{balanced}. A balanced family is one which satisfies, for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

@@

This techique can be used to turn coherent error into incoherent error therefore correlated error into uncorrelated error. One benefit of this is that non-Markovian noise will be made to be more Markovian. This is particularly useful because many routines exist that can assess the quality of a gateset, however in the presence of non-Markovianity most of them become unreliable.

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome \cite{Merkel2013}, and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs change the error to a depolorizing channel, the correlations in the noise can be reduced, and the process can be made much closer to Markovian, and as we will show, this comes at no real cost to the fidelity of the implemented gate. Finally, unlike in \cite{Wallman2016, Campbell2017, Ware2018} we note that this method of coherent error mitigation is cheap to implement - it only requires a random choice of pulse definition at run time, in addition to the extra storage required to store the pulse shapes - no classical preprocessing or runtime frame tracking is required.

@@

\section{Experimental Results}\label{experimental}
To demonstrate the realizability of this routine, we implemented it on a superconducting qubit, to calibrate an $RX(\frac{\pi}{2})$ pulse. The qubit being used had a 20.5 $\mu$s T1, and a 11.00 $\mu$s T2, and the pulse shapes being considered were 30ns long top hat pulse sequences, generated via a set-and-hold signal generator using the amplitudes shown in
